Discrete Quantum Theories and Computing

### 2018-03-18 dissertation.tex
dissertation.lyx "3.4.1 Complexified Finite Fields" Updating...

\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\mathbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\bf #1}}
\newcommand{\mathReal}{\mathbb{R}}
\newcommand{\mathComplex}{\mathbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\mathbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\CP}[1]{\mathComplex\mathbf{P}^{#1}}  % Complex projective space
\newcommand{\DCP}[1]{\mathbf{D}\mathComplex\mathbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\fnorm}[1]{\mathsf{N}(#1)}

@@

Our next objective is to develop more realistic discrete quantum theory
variants that exclude ``supernatural'' algorithms such as the one
presented above. Our first such plausible framework~\cite{HOSW2011}
is based on complexifiable finite fields. To incorporate complex numbers
for quantum amplitudes, we exploit the fact that the polynomial $x^{2}+1=0$
is \emph{irreducible} (has no solution) over a prime field $\ff{p}$
with $p$ odd if and only if $p$ is of the form~$4\ell+3$, with
$\ell$ a non-negative integer. In other words, $x^{2}+1=0$ is irreducible
over $\ff{3},\ff{7},\ff{11},\ff{19},\ldots$. We achieve our goal
by observing that any field in this family is extensible to a field~$\ff{p^{2}}$
whose elements can be viewed as discrete complex numbers with the
real and imaginary parts in~$\ff{p}$. In the field~$\ff{p^{2}}$,
the Frobenius automorphism of an element $\alpha$ (defined as $\alpha^{p}$)
represents the usual definition of complex conjugation~\cite{fieldexample}.

In a complexified finite field $\Fpp$, the Frobenius automorphism
that maps $\alpha\in\Fpp$ to $\alpha^{p}\in\Fpp$ acts like complex
conjugation. For example, in $\Fppx{3}$, we have $(2+i)^{3}=8+12i-6-i=2+11i$
which, in the field, is equal to $2-i$ since $11\equiv-1\pmod{3}$.

We define the \emph{field norm} $\fnorm{\cdot}$ as the map from $(a+ib)\in\Fpp$
to $a^{2}+b^{2}\in\Fp$, 
\begin{equation}
\fnorm{\alpha=a+ib}=a^{2}+b^{2}\ .\label{fieldnorm.eq}
\end{equation}
We avoid the square root in the discrete field framework because,
unlike the continuous case, the square root does not always exist.

>>>>

\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\mathbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\bf #1}}
\newcommand{\mathReal}{\mathbb{R}}
\newcommand{\mathComplex}{\mathbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\mathbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\CP}[1]{\mathComplex\mathbf{P}^{#1}}  % Complex projective space
\newcommand{\DCP}[1]{\mathbf{D}\mathComplex\mathbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\fnorm}[1]{\mathsf{N}\left({#1}\right)}

@@

Our next objective is to develop more realistic discrete quantum theory
variants that exclude ``supernatural'' algorithms such as the one
presented above. Our first such plausible framework~\cite{HOSW2011}
is based on complexifiable finite fields. To incorporate complex numbers
for quantum amplitudes, we exploit the fact that the polynomial $x^{2}+1=0$
is \emph{irreducible} (has no solution) over a prime field $\ff{p}$
with $p$ odd if and only if $p$ is of the form $4\ell+3$, with
$\ell$ a non-negative integer. In other words, $x^{2}+1=0$ is irreducible
over $\ff{3}$, $\ff{7}$, $\ff{11}$, $\ff{19}$, \ldots{}. We achieve
our goal by observing that any field in this family is extensible
to a field~$\ff{p^{2}}$ whose elements can be viewed as discrete
complex numbers with the real and imaginary parts in~$\ff{p}$. In
a complexified finite field $\Fpp$, the Frobenius automorphism that
maps $\alpha\in\Fpp$ to $\alpha^{p}\in\Fpp$ acts like complex conjugation
\cite{Artin1991,grove2002classical,DummitFoote2004}. For example,
in $\Fppx{3}$, we have $\left(2+\rmi\right)^{3}=8+12\rmi-6-\rmi=2+11\rmi$
which, in the field, is equal to $2-\rmi$ since $11\equiv-1\pmod{3}$.

We define the \emph{field norm} $\fnorm{\cdot}$ as the map from $(a+ib)\in\Fpp$
to $a^{2}+b^{2}\in\Fp$, 
\begin{equation}
\fnorm{\alpha=a+\rmi b}=a^{2}+b^{2}\,.\label{fieldnorm.eq}
\end{equation}
We avoid the square root in the discrete field framework because,
unlike the continuous case, the square root does not always exist.
